On the discovery of K-enhanced and Mg-depleted stars throughout the Milky Way

### 2018-02-08 article/ms.tex
worked in graphical changes, included the 0.2 quadrature flooring in both table and plots for LAMOST abundance errrors

\begin{abstract}
Stars with unusual elemental abundances offer clues about rare astrophysical events or nucleosynthetic pathways. Stars with significantly depleted magnesium and enhanced potassium (e.g., $[{\rm K}/{\rm Fe}] > 1$; $[{\rm Mg}/{\rm Fe}] < -0.5$) have to date only been found in the massive globular cluster NGC~2419, and to a lesser extent, NGC~2808. The origin of this abundance signature remains unknown, as does the reason for its apparent exclusivity to these two globular clusters. Here we present 112 field stars, identified from \LamostGiants\ \lamost\ giant stars \todo{using a data-driven model produced by \tc}, that show significantly enhanced [K/Fe] and possibly depleted [Mg/Fe] abundance ratios.
Our sample spans a full range of metallicities ($-1.5 < [{\rm Fe}/{\rm H}] < 0.3$), yet none show abundance ratios of [K/Fe] or [Mg/Fe] that are as extreme as those observed in NGC~2419. 
We conclude that the nucleosynthetic process producing the anomalous abundances in [K/Fe] and [Mg/Fe] probably (but not necessarily) occurs at a wide range of metallicities. This would suggest that pollution scenarios that are limited to early epochs (e.g., Population III supernovae) are an unlikely explanation. This sample is expected to help guide modelling \todo{and observational} attempts to explain the origin of the Mg-K abundance signature, which we show exists throughout the Milky Way.
\end{abstract}

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{posterchildof13.png}
    \caption{Pseudo-continuum-normalised \lamost\ spectra for the Mg-depleted and K-enhanced candidate J034458.82+592955.1. The data  are shown in black and the best-fitting data-driven model from \tc\ is shown in red. We show zoom-in axes around the magnesium triplet and potassium doublet, which we used to identify Mg-K candidates. Note that $[\alpha/{\rm Fe}]$ is a label in \tc\ model used, and J034458.82+592955.1 shows Mg depletions relative to the estimated [$\alpha$/Fe] value.}
    \label{posterchild}
\end{figure*}

\begin{figure}
	\includegraphics[width=\columnwidth]{histof113.png}
    \caption{Metallicity distribution for all 112 candidate stars identified with K-enhancement and Mg-depletion.}
    \label{mhist}
\end{figure}

\begin{figure}
	\includegraphics[width=\columnwidth]{Kvsmgmod.png}
    \caption{Comparison of [K/Fe] vs [$\alpha$/Fe] for the candidate stars with [K/Fe] vs [Mg/Fe] abundances for NGC 2419 and NGC 2808 \citep{cohenkirby2012, mucciarelli2012, mucciarelli2015}. See text for discussion on the use of [$\alpha$/Fe] as a proxy for [Mg/Fe]. The dotted line is arbitrary, and is present simply as a visual aide.}
    \label{KvsMg}
\end{figure}

@@

\label{sec:scenarios}
\begin{table*}
\centering
\caption{Tabulated data for 10 of the 112 candidate stars. Data for the full sample is available online.}
\label{data}
\begin{tabular}{cccccccccccccc}
\hline
\textbf{2MASSID} & \textbf{RA} & \textbf{DEC} & \textbf{S/N} & \textbf{V$_\textrm{r}$} & \textbf{T$_\textrm{eff}$} & $\boldsymbol \log{ \, \textbf{g}}$ & \textbf{{[}Fe/H{]}} & \textbf{{[}$\alpha$/Fe{]}} & \textbf{$\boldsymbol \chi_r ^\textbf{2}$} & \textbf{{[}K/Fe{]}} & \textbf{Error} & \textbf{{[}Na/Fe{]}} & \textbf{Error} \\ 
- & {[}deg{]} & {[}deg{]} & [pixel$^{-1}]$ & {[${\rm km\,s}^{-1}$]} & {[}K{]} & $[{\rm cm\,s}^{-2}]$ & {[}dex{]} & {[}dex{]} & - & {[}dex{]} & {[}dex{]} & {[}dex{]} & {[}dex{]} \\ \hline
J000908.89+124821.9 & 2.29 & 12.81 & 10 & -78.55 & 4892 & 2.82 & -0.12 & 0.33 & 0.19 & 1.08 & 0.16 & 0.86 & 0.12 \\
J002619.36+565612.1 & 6.58 & 56.94 & 31 & -39.57 & 4205 & 1.49 & 0.22 & 0.06 & 0.76 & 0.62 & 0.08 & 0.39 & 0.10 \\
J002907.61+354701.0 & 7.28 & 35.78 & 15 & -82.74 & 5116 & 3.56 & 0.24 & 0.17 & 0.18 & 1.01 & 0.06 & -0.20 & 0.00 \\
J003451.10+424543.0 & 8.71 & 42.76 & 15 & -79.15 & 4460 & 2.10 & -0.56 & 0.24 & 0.31 & 0.96 & 0.26 & -0.16 & 0.00 \\
J005649.98+391722.9 & 14.21 & 39.29 & 41 & -46.77 & 4578 & 2.67 & 0.31 & 0.10 & 0.44 & 0.76 & 0.11 & -0.02 & 0.00 \\
J010305.95+043445.9 & 15.77 & 4.58 & 90 & 20.69 & 4646 & 2.72 & 0.10 & 0.05 & 1.17 & 0.95 & 0.06 & -0.07 & 0.00 \\
J011949.36+063411.4 & 19.96 & 6.57 & 479 & -28.18 & 4820 & 2.61 & -0.29 & 0.09 & 2.38 & 1.03 & 0.00 & -0.04 & 0.00 \\
J013039.08+404843.8 & 22.66 & 40.81 & 71 & -80.34 & 4355 & 2.01 & -0.09 & 0.14 & 0.93 & 0.92 & 0.05 & 0.46 & 0.05 \\
J030101.42+560042.3 & 45.26 & 56.01 & 88 & -39.27 & 4878 & 2.53 & -0.32 & 0.09 & 0.67 & 1.00 & 8.70 & -0.02 & 0.00 \\
J032423.46+425429.6 & 51.10 & 42.91 & 23 & -56.74 & 4649 & 2.63 & -0.11 & -0.01 & 0.34 & 1.12 & 0.09 & 0.80 & 0.04 \\ \hline
\end{tabular}
\end{table*}

@@

We find extreme abundance ratios of [K/Fe] $\approx 1.5$ \todo{can we say they were aprox 1.5? only 1 star was that high... surely an 'up to' or an 'as high as' is more accurate} in our sample, but none that exceed the levels observed in NGC 2419 (up to [K/Fe] $\approx$ 2). While our analysis suggests the presence of substantial [Mg/Fe] depletion, follow-up observations are necessary to quantify the level of Mg-depletion. The spread in metallicities among our sample suggests that the mechanism responsible for the Mg-K abundance signature is not limited to early epochs (e.g., Population III supernova) or low-metallicity environments. We highlight SAGB-binaries as a plausible polluter candidate for the Mg-K abundance signature \todo{within the disk}. We encourage future high-resolution observations and modelling efforts to understand this peculiar nucleosynthetic pathway.

>>>>

\begin{abstract}
Stars with unusual elemental abundances offer clues about rare astrophysical events or nucleosynthetic pathways. Stars with significantly depleted magnesium and enhanced potassium (e.g., $[{\rm K}/{\rm Fe}] > 1$; $[{\rm Mg}/{\rm Fe}] < -0.5$) have to date only been found in the massive globular cluster NGC~2419, and to a lesser extent, NGC~2808. The origin of this abundance signature remains unknown, as does the reason for its apparent exclusivity to these two globular clusters. Here we present 112 field stars, identified from \LamostGiants\ \lamost\ giant stars using a data-driven model produced by \tc, that show significantly enhanced [K/Fe] and possibly depleted [Mg/Fe] abundance ratios.
Our sample spans a full range of metallicities ($-1.5 < [{\rm Fe}/{\rm H}] < 0.3$), yet none show abundance ratios of [K/Fe] or [Mg/Fe] that are as extreme as those observed in NGC~2419. 
We conclude that the nucleosynthetic process producing the anomalous abundances in [K/Fe] and [Mg/Fe] probably (but not necessarily) occurs at a wide range of metallicities. This would suggest that pollution scenarios that are limited to early epochs (e.g., Population III supernovae) are an unlikely explanation. This sample is expected to help guide modelling  attempts to explain the origin of the Mg-K abundance signature, which we show exists throughout the Milky Way.
\end{abstract}

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{posterchild.png}
    \caption{Pseudo-continuum-normalised \lamost\ spectra for the Mg-depleted and K-enhanced candidate J034458.82+592955.1. The data  are shown in black and the best-fitting data-driven model from \tc\ is shown in red. We show zoom-in axes around the magnesium triplet and potassium doublet, which we used to identify Mg-K candidates. Note that $[\alpha/{\rm Fe}]$ is a label in \tc\ model used, and J034458.82+592955.1 shows Mg depletions relative to the estimated [$\alpha$/Fe] value.}
    \label{posterchild}
\end{figure*}

\begin{figure}
	\includegraphics[width=\columnwidth]{mhist.png}
    \caption{Metallicity distribution for all 112 candidate stars identified with K-enhancement and Mg-depletion.}
    \label{mhist}
\end{figure}

\begin{figure}
	\includegraphics[width=\columnwidth]{KvsMg.png}
    \caption{Comparison of [K/Fe] vs [$\alpha$/Fe] for the candidate stars with [K/Fe] vs [Mg/Fe] abundances for NGC 2419 and NGC 2808 \citep{cohenkirby2012, mucciarelli2012, mucciarelli2015}. See text for discussion on the use of [$\alpha$/Fe] as a proxy for [Mg/Fe]. The dotted line is arbitrary, and is present simply as a visual aide.}
    \label{KvsMg}
\end{figure}

@@

\label{sec:scenarios}
\begin{table*}
\centering
\caption{Tabulated data for 10 of the 112 candidate stars. Data for the full sample is available online.}
\label{data}
\begin{tabular}{cccccccccccccc}
\hline
\textbf{2MASSID} & \textbf{RA} & \textbf{DEC} & \textbf{S/N} & \textbf{V$_\textrm{r}$} & \textbf{T$_\textrm{eff}$} & $\boldsymbol \log{ \, \textbf{g}}$ & \textbf{{[}Fe/H{]}} & \textbf{{[}$\alpha$/Fe{]}} & \textbf{$\boldsymbol \chi_r ^\textbf{2}$} & \textbf{{[}K/Fe{]}} & \textbf{Error} & \textbf{{[}Na/Fe{]}} & \textbf{Error} \\ 
- & {[}deg{]} & {[}deg{]} & [pixel$^{-1}]$ & {[${\rm km\,s}^{-1}$]} & {[}K{]} & $[{\rm cm\,s}^{-2}]$ & {[}dex{]} & {[}dex{]} & - & {[}dex{]} & {[}dex{]} & {[}dex{]} & {[}dex{]} \\ \hline
J000908.89+124821.9 & 2.29 & 12.81 & 10 & -78.55 & 4892 & 2.82 & -0.12 & 0.33 & 0.19 & 1.08 & 0.26 & 0.86 & 0.23 \\
J002619.36+565612.1 & 6.58 & 56.94 & 31 & -39.57 & 4205 & 1.49 & 0.22 & 0.06 & 0.76 & 0.62 & 0.22 & 0.39 & 0.22 \\
J002907.61+354701.0 & 7.28 & 35.78 & 15 & -82.74 & 5116 & 3.56 & 0.24 & 0.17 & 0.18 & 1.01 & 0.21 & -0.20 & 0.20 \\
J003451.10+424543.0 & 8.71 & 42.76 & 15 & -79.15 & 4460 & 2.1 & -0.56 & 0.24 & 0.31 & 0.96 & 0.33 & -0.16 & 0.20 \\
J005649.98+391722.9 & 14.21 & 39.29 & 41 & -46.77 & 4578 & 2.67 & 0.31 & 0.1 & 0.44 & 0.76 & 0.23 & -0.02 & 0.20 \\
J010305.95+043445.9 & 15.77 & 4.58 & 90 & 20.69 & 4646 & 2.72 & 0.1 & 0.05 & 1.17 & 0.95 & 0.21 & -0.07 & 0.20 \\
J011949.36+063411.4 & 19.96 & 6.57 & 479 & -28.18 & 4820 & 2.61 & -0.29 & 0.09 & 2.38 & 1.03 & 0.20 & -0.04 & 0.20 \\
J013039.08+404843.8 & 22.66 & 40.81 & 71 & -80.34 & 4355 & 2.01 & -0.09 & 0.14 & 0.93 & 0.92 & 0.21 & 0.46 & 0.21 \\
J030101.42+560042.3 & 45.26 & 56.01 & 88 & -39.27 & 4878 & 2.53 & -0.32 & 0.09 & 0.67 & 1 & 8.70 & -0.02 & 0.20 \\
J032423.46+425429.6 & 51.1 & 42.91 & 23 & -56.74 & 4649 & 2.63 & -0.11 & -0.01 & 0.34 & 1.12 & 0.22 & 0.80 & 0.20 \\ \hline
\end{tabular}
\end{table*}

@@

We find extreme abundance ratios of up to [K/Fe] $\approx 1.5$ in our sample, but none that exceed the levels observed in NGC 2419 (up to [K/Fe] $\approx$ 2). While our analysis suggests the presence of substantial [Mg/Fe] depletion, follow-up observations are necessary to quantify the level of Mg-depletion. The spread in metallicities among our sample suggests that the mechanism responsible for the Mg-K abundance signature is not limited to early epochs (e.g., Population III supernova) or low-metallicity environments. We highlight SAGB-binaries as a plausible polluter candidate for the Mg-K abundance signature within the disk. We encourage future high-resolution observations and modelling efforts to understand this peculiar nucleosynthetic pathway.
